\chapter*{\Large Acknowledgments}
% \enlargethispage{2\baselineskip} \Large

First, I would express my deepest gratitude to my supervisor, Professor Roberto Togneri. His coursework unit is the most clearly one I got during my Bachelor study, and he shows me the light to step-in this journey. With many times of my wasting time consultant, he always kind, and never complained once. Same gratitude to my supervisor Professor Defeng (David) Huang. Since Roberto's retirement, and even thought our topic is not such closed, he read what I wrote in-depth, undoubtly a expertise and professional. In Chinese language, we said the Teachers is like our fathers, and this is my ultimately lucky to have both of them.

Same deepest gratitude goes to my family and relatives. My parents, Weihu He and Weili Chi, what they give me can never pay back. I got many happiness from my sister, Min Hu, and her families: Wen He, Jiahui, Jiaxuan, Jennie, and Jiacheng. Talking with them is the most relaxable time for me. The same warmth goes to my aunt, Xinyun He, who has been like a second mom to me, to her husband, and to my elder brother, Jie Hu, who brought me to university. I share my happiness with my younger brother, Zhanpeng He, who is ongoing graduate and being a professional writer, and my warmest uncle, Weida He, all along his family. I have being overseas too long, and families left along my way. I hope they are well in the heaven.

A special thanks to my partner, Hanyu Meng. She is take over my boring life, and make every day colorful. We have the same topic, similar research interest, and we can always have a deep discussion and make every day joyful and productive. I am so lucky to have her, and I will always be grateful for her support and love.

I would also like to thank my colleagues and friends in UWA, who have been a great support system throughout my PhD journey. Yuliang. Zhihui and Kaibu...

Lastly, my sincere appreciation goes to the Graduate Research School and the EECE administration, Leila Easter, Lea Towler, along with many other warm and supportive members. Their responses were always prompt, helpful, and kind. Staying in UWA has felt like a family to me, and I will always remember and miss my time here. This research was carried out while the author was in receipt of a University of Western Australia International Fee Scholarship (UIFS) and a University Postgraduate Award (UPA). I sincerely thank the University for this financial support, which allowed me to pursue research topics that I found interesting without the financial pressure. I am also grateful to UWA High Performance Computing and IT (HPC-IT) for their strong support. The team consistently provided wholehearted assistance to our research, which was essential in completing my PhD.